---
title: "MLB 2017"
author: "Jensen Holm & Kyle Knapp"
date: "2022-11-07"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Libraries
```{r include=TRUE, message=FALSE, warning=FALSE}
library(grid)
library(tidyverse)
library(lubridate)
library(boot)
```

# Read in the data
```{r include = TRUE}
# should work with any computer as long as you have
# the data folder to go with this
dataFilePath <- str_c(getwd(), "/data/", sep = "")

# function to make importing data into this project
# seamless
importCSV <- function(fileName) {
  read.csv(str_c(dataFilePath, str_c(fileName, ".csv", sep = ""), sep = ""))
}
```


```{r include = TRUE}
statCast <- importCSV("statcast2017")
bbref <- importCSV("bbref2017")
trackman <- importCSV("trackman")
```


# Joining the two datasets together with dplyr and stringr

Joining the statcast dataset and the baseball reference
datasets together on the players name using stringr and a left join
```{r include=TRUE}
statCastName <- statCast %>% mutate(
  Name = str_c(
    str_replace_all(first_name, " ", ""), 
    str_replace_all(last_name, " ", ""), 
    sep = " "
    )
)  

bsbl <- left_join(
  statCastName, 
  bbref, 
  by = "Name"
)
```


# DATA VISUALIZATION
```{r include = TRUE, warning=FALSE, message=FALSE}
# without applying this filter on the data set, there were
# a lot of players with too few plate appearances messing 
bsblSub <- bsbl %>% subset(attempts > 100)  

bsblSub %>% ggplot(aes(
  x = anglesweetspotpercent, 
  y = avg_distance, 
  color = attempts, 
  alpha = 0.5
  )) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Average Hit Distance as a funciton of launch angle sweet spot %" ,
    x = "Launch Angle Sweet Spot %",
    y = "Average Hit Distance (ft.)"
  )
```

```{r, include=TRUE, message=FALSE, warning=FALSE}
# heatmap of batting average compared to avg_hit_speed and avg_laungh_angle
bsblSub %>% ggplot(aes(
  x = round(avg_hit_angle, digits=0),
  y = round(avg_hit_speed, digits=0),
  fill = BA
)) + 
  geom_tile() + 
  labs(
    title = "Average Hit Speed vs Average Hit angle grouped by BA",
    x = "Average Hit Angle",
    y = "Average Hit Speed"
  )
```

The above heatmap is not amazing, I had to round the numbers for average hit
speed and average hit angle down to an integer because without it it was no 
longer a heatmap, but a scatterplot with funny looking points. Still I think
that this illustrates that having a higher average hit angle and hit speed than
everyone else does not mean that you have a higher batting average. The sweet 
spot appears to be in the middle of the heatmap as that is where the colors
appear to be lighter in color. Thus indicating a higher batting average

```{r, include=TRUE, message=FALSE, warning=FALSE}
bsblSub %>% ggplot(aes(x = HR, y = RBI)) + geom_density_2d_filled() +
  labs(
    title = "Runs Batted In as a function of Home Runs in 2017 (Density Plot)"
  )
```


#### making strikezone visualization with trackman dataset
```{r include=TRUE}
# constanst for the strikezone were decided based on this
# article -> https://www.baseballprospectus.com/news/article/40891/prospectus-feature-the-universal-strike-zone/

in2ft <- function(inches) {
  inches / 12
}

leftSZ <- in2ft(-(19.94 / 2)) * 2
rightSZ <- in2ft(19.94 / 2) * 2
heightSZ <- in2ft(44.08)
minySZ <- in2ft(18.29)
```

According to baseballprospectus.com, the dimensions for the universal
strikezone are as follows:

Height :`r heightSZ` ft.
Width : `r rightSZ` ft.

#### Function for creating a strikezone plot given a trackman dataset
```{r include=TRUE}
strikeZonePlot <- function(dataset) {
  dataset %>% ggplot(
    aes(
      x = PlateLocSide,
      y = PlateLocHeight,
      color = TaggedPitchType
    ) 
  ) + geom_point() + 
    geom_rect(
      xmin = leftSZ,
      xmax = rightSZ,
      ymin = minySZ,
      ymax = heightSZ,
      alpha = 0,
      color = "black"
    ) + 
    scale_x_continuous(limits = c(-5, 5)) +
    scale_y_continuous(limits = c(-1, 6))
}
```


#### Strike Zone graph filtered with only pitches thrown from Right Handed Pitchers
```{r include=TRUE, message=FALSE, warning=FALSE}
trackman %>% subset(PitcherThrows == "Right") %>% strikeZonePlot() +
  labs(
    title = "Pitches thrown by right handed pitchers colored by pitch type "
  )
```


#### Strike Zone graph filtered with only pitches thrown from Left Handed Pitchers
```{r include=TRUE, message=FALSE, warning=FALSE}
trackman %>% subset(PitcherThrows == "Left") %>% strikeZonePlot() +
  labs(
    title = "Pitches thrown by left handed pitchers colored by pitch type"
  )
```



# T-Test
#### Comparing average hit speed between the colorado rockies <br>
#### and the rest of Major League Baseball
```{r}
colorado <- bsbl %>% subset(Team == "Colorado")
everyoneElse <- bsbl %>% subset(Team != "Colorado")

t.test(
  colorado$avg_hit_speed, 
  everyoneElse$avg_hit_speed
)
```

This t-test is interesting because since the Rockies play at 5,000+ ft. <br>
of elevation, the baseball travels further and faster there. However with the <br>
t-test we conclude that there is a difference between the two groups, but I <br>
expected the rockies to be higher. Then again, they play just half of their games <br>
at their home stadium, and play in the AL West which is arguably the toughest <br>
division in baseball. Espicially in 2017.

# BOOTSTRAPPING

#### Functions to perform Non-Parametric-Bootstrapping
```{r include=TRUE}

npBoots <- function(data_set, num_samples, func=mean) {
  statFunc <- function(x, i){func(x[i])}
  return (
    boot(c(data_set), statFunc, num_samples)
  )
}

```

#### Bootstrapping 100 samples for average hit angle in 2017 
```{r include=TRUE}
npBoots(bsblSub$avg_hit_angle, 100)
```


#### Functions related to performing parametric bootstrapping
```{r include=TRUE}

paramBoots <- function(data_set, num_samples) {
  pBoots <- matrix(
    NA, nrow = length(data_set), ncol = num_samples
  )
  for (i in 1:num_samples) {
    pBoots[, i] <- rnorm(
      n = length(data_set),
      mean = mean(data_set),
      sd = sd(data_set)
    )
  }
  return (pBoots)
}


# for the func argument in this function, pass in a function
# 'mean' or 'median'
paramBootsStat <- function(func, parameterBoots) {
  pbs <- vector(length = ncol(parameterBoots))
  for (i in 1:ncol(parameterBoots)) {
    pbs[i] <- func(parameterBoots[,i])
  }
  return (pbs)
}

```

Using the above functions to evaluate the standart deviation of <br>
the bootstrap sample mean, and median for season total hits in 2017

```{r include=TRUE}
b <- paramBoots(bsbl$H,100)
bootMedians <- sd(paramBootsStat(median, b))
bootMeans <- sd(paramBootsStat(mean, b))
```

Standard deviation of bootstrapped medians -> `r bootMedians` <br>
Standart deviation of bootstrapped means -> `r bootMeans` <br>


# Pivoting data into wide format

<br>
This would create a new dataset with columns for each unique value in the Team column, <br> 
and the values for barrels per plate appearance <br> 
column would be spread across these new columns. <br>
<br>
```{r include=TRUE,  R.options=list(max.print=10)}
bsblSub %>% pivot_wider(
  names_from = Team, 
  values_from = brl_pa
)
```

# utilizing lubridate functions with our data
```{r include=TRUE}
# creating the dates vector
dates <- trackman$Date %>% month()
# adding it to our dataframe
trackman$month = month.abb[dates]
```

The trackman data that we have is from one game. <br>
This game was played in the month of `r trackman$month[1]`

# Data Dictionary
```{r indluce=FALSE}

dataDictionary <- tibble(Variable = colnames(bsbl),
                         Description = c("Last name of player",
                                         "First name of player",
                                         "Player identification number",
                                         "Number of batting attempts",
                                         "Average launch angle of the ball off the bat (degrees)",
                                         "Percentage of hits where the launch angle of the ball fell within the sweet spot",
                                         "Maximum ball speed of a players single hit (mph)",
                                         "Average ball speed of a player's hits (mph)",
                                         "Fly ball line drive percentage",
                                         "Ground ball percentage",
                                         "Maximum distance of a ball hit (feet)",
                                         "Average distance of a ball hit (feet)",
                                         "Average distance of a home run (feet)",
                                         "ev95Plus",
                                         "ev95percent",
                                         "Number of barrels hit in a season",
                                         "Barrel percentage",
                                         "Barrels per plate appearance",
                                         "Player's full name",
                                         "Player's ID number",
                                         "Player's baseball reference ID number",
                                         "Season year",
                                         "Player's age",
                                         "Level of play",
                                         "Team player is on",
                                         "Games played",
                                         "Plate appearences",
                                         "At-bats",
                                         "Runs scored",
                                         "Number of hits",
                                         "Singles",
                                         "Doubles",
                                         "Triples",
                                         "Home runs",
                                         "Runs batted in",
                                         "Walks",
                                         "Intentional walks",
                                         "Unintentional walks",
                                         "Strikeouts",
                                         "Hit by pitch",
                                         "Shutouts",
                                         "Sacrifice Flies",
                                         "Ground into double play",
                                         "Stolen bases",
                                         "Caught Stealing",
                                         "Batting average",
                                         "On base percenatage",
                                         "Slugging percentage",
                                         "On base plus slugging percentage"),
                         Type = map_chr(bsbl, .f = function(x){typeof(x)[1]}),
                         Class = map_chr(bsbl, .f = function(x){class(x)[1]}))

# printing it out nicely
knitr::kable(dataDictionary)
```




